### 2022-02-21 manuscript/tab_rma_template.Rmd
update 2022-02-21 19:22:02

```{r tabref{{i}}, results = 'asis'}
p <- processes[{{i}}]
    tb <- modres[[{{i}}]]$refcat
    # knitr::kable(tb, caption = paste0("Reference category for each variable for process ", p))
    datatable(
          tb,
          rownames = FALSE,
          caption = paste0("Category not compared to the grand mean for each variable for process ", p))
```

```{r tabrma{{i}}, results = 'asis'}
p <- processes[{{i}}]
    tb <- modres[[{{i}}]]$rma
    tb$format <- as.integer(tb$p < .05)
    formatStyle(
      formatRound(
        datatable(
          tb,
          rownames = FALSE,
          caption = paste0("Meta-regression coefficients for ", p),
          options = list(columnDefs = list(list(
            targets = 6, visible = FALSE
          )),
          "pageLength" = nrow(tb))
        ),
        columns = c('Estimate', "se", 'p', "VIF"),
        digits = 2
      ),
      "format",
      target = "row",
      backgroundColor = styleEqual(c(0, 1), c('gray', 'white'))
    )
```